---
title: "FVCM Package"
output: github_document
---

# FVCM (The Floodplain Vegetation Condition model)

This package was created to simulate floodplain vegetation condition responses to river inundation over time. I developed a state and transition simulation model (STSM) using rasterised layers of inundation threshold, vegetation community type and condition. The STSM approach provides dynamic time-series of vegetation condition in response to different inundation sequences (defined in terms of frequency, timing and duration) for individual pixels in the vegetation rasters. The implementation of the FVCM is a major improvement to typical vegetation condition assessments and attendant decision making, providing a spatially and temporally explicit framework for quantifying vegetation outcomes in ways that have not previously been possible.

Combining the STSM with hydrologic time-series and associated inundation sequences allows the projection of the initial vegetation state (veg type and veg condition) across a hydrologic time series representing different flow scenarios. Outputs from the model consist of annual rasters describing vegetation type and condition within each pixel across the floodplain area. Individual raster extents were aligned with zones in the inundation models.

## Installation

Clone this repository and install using `devtools`:

```{r}

# install devtools package
install.packages(c("devtools"))

devtools::install_github("LukeMcPhan/FVCM")

# Remove the package zip after installation
unlink("FVCM.zip")

# load library
# library(FVCM)

```

## Package Workflow

This package contains five functions that are reflective of different analyses and must be used in the following order for analysis to run: 

1) `prepare_data()` – Loads and processes vegetation, flow, and inundation data.
2) `format_rules()` - define system states and transition rules. Formats the transition rules for analysis.
3) `spatial_preprocessing()` - Splits and rasterizes spatial data for analysis.
4) `spells_analysis()` - Analyzes flow data and calculates applicable transitions.
5) `matrix_projection()` - Projects vegetation state transitions spatially.
6) `save_output()` - Saves annual rasters and summaries.

## Example Usage

```{r}
# Load required packages
library(FVCM)

# Prepare input data
data <- prepare_data(
  transition_rules_file = "data/rules.csv",
  flow_files = list("data/flow1.csv", "data/flow2.csv"),
  inundation_raster_path = "data/inundation.tif",
  vegetation_gdb_path = "data/vegetation.gdb",
  vegetation_layer_name = "RRGF"
)

# Format rules
rules <- format_rules(data$transition_rules, data$ordered_states)

# Preprocess spatial data
spatial <- spatial_preprocessing(data$inundation_raster, data$vegetation_data)

# Analyze spells
spell_output <- spells_analysis(
  flow_data = data$flow_data[[1]],
  inundation_raster = data$inundation_raster,
  between_spell_lengths = 30,
  water_year_start = 4,
  save_spell_rule_timeseries = TRUE
)

# Project matrix transitions
projection <- matrix_projection(spatial, spell_output, rules)

# Save outputs
save_output(
  matrix_projection_output = projection,
  rule_set_name = "basecase",
  vegetation_layer_name = "RRGF",
  inundation_raster_name = "inundation",
  flow_data_file_name = "flow1",
  scenario_name = "test",
  water_years = spell_output$water_years,
  inundation_timeseries = spell_output$inundation_rule_timeseries
)
```

## Authors

Developed by [Luke McPhan], [CSIRO].

## License

This package is released under the MIT License.
